---
title: "NCREIF Returns Exploration"
author: "Nicolás Cuesta Quintero"
date: "3/13/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. NCREIF Highlights

Quarterly returns calculated as

$$r_{t} = y_{t}+g_{t}.$$

where $y$ is the income return and $g$ is the growth or appreciation return, given by

$$y_{t}=\frac{NOI_{t}}{BMV_{t}-\frac{1}{2}\left(PS_{t}-CI_{t}\right)-\frac{1}{3}NOI_{t}},$$
$$g_{t}=\frac{EMV_{t}-BMV_{t}+\left(PS_{t}-CI_{t}\right)}{BMV_{t}-\frac{1}{2}\left(PS_{t}-CI_{t}\right)-\frac{1}{3}NOI_{t}}.$$

# 1. Data import

Upload required packages.

```{r Import packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(readxl)
library(lubridate)
library(magrittr)
```

Read required data. Start the analysis with only property types

```{r Import NCREIF returns}
ncf_source <- "../Source/SourceFiles/NCREIF/NPI Returns 3-9-21.xlsx"

map_dfr(c("NPI - National", "NPI - Property Type"),
        function(x) {
          read_xlsx(path = ncf_source, sheet = x)
         }) %>% 
  transmute(
    Date = yq(str_c(Year, "0", Quarter)) %m+% months(3) %m+% days(-1),
    PropertyType = factor(replace_na(PropertyType, "All"), levels = c("All", "A", "H", "I", "O", "R")),
    y = `Income Return`,
    g = `Capital Return`,
    r = `Total Return`
  ) -> ncf_dataset

glimpse(ncf_dataset)
```

# 2. Exploratory data analysis

## 2.1. Descriptive statistics


Positive serial-correlation

```{r}
returns <- ts(
    filter(ncf_dataset, PropertyType == "All", Date <= as.Date("2020-03-31"))$r, 
    start = 1978, 
    frequency = 4
  )

forecast::ggtsdisplay(
  returns, 
  lag.max = 30, 
  plot.type = "scatter", 
  theme = theme_bw(),
  ylab = "NPI Total Returns", 
  xlab = "Year"
)
```

- Hospitality starts 18 quarters late.
- No explicit data gaps.
- Income returns seem to be uniform.
- Except for Hospitality, sample does not show negative NOI.
- Capital returns means are lower but standard deviations are higher.

```{r Descriptive statistics}
ncf_dataset %>% 
  group_by(PropertyType) %>% 
  skimr::skim()
```

- There is clearly no winner or loser. All property types seems to have similar mean.
- Hispitality has clearly more outliers.
- Income returns are less volatile than capital returns.

```{r Returns boxplot per index}
ncf_datasetp <- ncf_dataset %>% 
  pivot_longer(cols = c("y", "g", "r"), names_to = "Index", values_to = "Return") %>% 
  mutate(Index = factor(Index, levels = c("y", "g", "r"))) 

ncf_datasetp %>% 
  ggplot(mapping = aes(x = Index, y = Return)) +
  geom_boxplot() +
  facet_wrap(~PropertyType, nrow = 1) +
  scale_y_continuous(labels = percent) +
  theme_bw()
```

- Clearly income return is greater than 0 and shifts positively the capital return to get the total one.
- Income appear to by leptokurtic.

```{r Returns shape per index}
ncf_datasetp %>% 
  ggplot(mapping = aes(x = Return, fill = Index, color = Index)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~PropertyType, ncol = 2) +
  scale_x_continuous(labels = percent) +
  theme_bw()
```

```{r NPI decomposition, message=FALSE, warning=FALSE}
ncf_dataset %>% 
  filter(PropertyType == "All") %>% 
  select(r, g, y) %>% 
  GGally::ggpairs() +
  scale_x_continuous(labels = percent) +
  # scale_y_continuous(labels = percent) +
  theme_bw()
```

## 2.2. Returns decomposition

Now we want to know how well would Property Types be hedged by total index. Therefore, we bulild the same graphic using total capital return against property level capital returns.

```{r Property types vs. NPI capital returns, message=FALSE, warning=FALSE}
ncf_dataset %>% 
  select(Date, PropertyType, g) %>% 
  pivot_wider(names_from = PropertyType, values_from = g) %>% 
  select(-c("Date")) %>% 
  drop_na() %>%
  `colnames<-`(c("Nation", "Apartments", "Industrial", "Office", "Retail", "Hospitality")) %>% 
  GGally::ggpairs() +
  scale_x_continuous(labels = percent) +
  theme_bw() +
  theme(
    axis.text.x=element_text(angle = 90, vjust = 0.5, size = 0.75),
    strip.background = element_blank(),
    strip.placement = "outside"
  )
```

Let's repeat the exercise with income returns.

```{r Property types vs. NPI income returns, message=FALSE, warning=FALSE}
ncf_dataset %>% 
  select(Date, PropertyType, y) %>% 
  pivot_wider(names_from = PropertyType, values_from = y) %>% 
  select(-c("Date")) %>% 
  drop_na() %>% 
  GGally::ggpairs() +
  scale_x_continuous(labels = percent) +
  theme_bw() +
  theme(axis.text.x=element_text(angle = 45, vjust = 0.5))
```

## 2.3. Property types portfolios analysis

This is spot vs. spot analysis.

```{r Portfolios construction, cache = TRUE}
# Grid of all possible portfolios (No short-sales only real estates)
port_grid <- expand.grid(
  wA = seq(0, 1, 0.05),
  wI = seq(0, 1, 0.05),
  wO = seq(0, 1, 0.05),
  wR = seq(0, 1, 0.05),
  wH = seq(0, 1, 0.05)
) %>% 
  mutate(Filter = wA + wI + wO + wR + wH) %>%
  filter(Filter == 1) %>% 
  select(wA:wH) %>% 
  mutate(key = row_number()) %>% 
  as_tibble()

port_model <- function(wA, wI, wO, wR, wH) {

  # Modified NCREIF dataset
  df <- ncf_dataset %>% 
    select(Date, PropertyType, g) %>% 
    pivot_wider(names_from = PropertyType, values_from = g) %>% 
    drop_na()
  
  # Calculation portfolio-wise
  n <- length(wA)
  v <- vector("list", n)
  for (k in 1:n) {
    v[[k]] <- df %>%
      mutate(Portfolios = A * wA[k] + I * wI[k] + O * wO[k] + R * wR[k] + H * wH[k]) %>% 
      select(Date, NPI = All, Portfolios) %>% 
      mutate(Index = 100 * cumprod(1+ Portfolios),
             Delta = c(Index[1] - 100, diff(Index)))
  }
  
  # Results
  return(v)
}

# Generate returns for each portfolio
port_returns <- port_grid %>% 
  select(key, wA:wH) %>% 
  mutate(Portfolio = port_model(wA, wI, wO, wR, wH))

glimpse(port_returns)
```

The most hospitality, the less the correlation.

```{r Property types portfolios parallel lines, message=FALSE, warning=FALSE}
# Parallel lines technique
port_returns %>% 
  mutate(rho = unlist(map(Portfolio, function(x) cor(x$NPI, x$Portfolios)))) %>% 
  GGally::ggparcoord(columns = 2:6, groupColumn = 8, scale = "globalminmax", showPoints = TRUE) +
  theme_bw()
```

## 2.4. Financials

```{r Import NCREIF financials}
map_dfr(c("NPI - National", "NPI - Property Type"),
        function(x) {
          read_xlsx(path = ncf_source, sheet = x)
         }) %>% 
  transmute(
    Date = yq(str_c(Year, "0", Quarter)) %m+% months(3) %m+% days(-1),
    PropertyType = factor(replace_na(PropertyType, "All"), levels = c("All", "A", "H", "I", "O", "R")),
    NOI,
    MV,
    BMV = MVLag1
  ) -> ncf_financials

glimpse(ncf_financials)
```

Apartments, Office and Industrial has a huge correlation with the national index. We want to analyze, histrically, what was their weight on the NPI.

- Office has a huge weight on the index, so correlation mught be spurious
- Apartments and industrial has relatively low participation and their correlation is relaly huge.
- Retail is an odd case.
- Hospitality might not be well represented by the index.

```{r}
ncf_fplot <- ncf_financials %>% 
  transmute(
    Date = Date, 
    PropertyType = PropertyType,
    Variable = MV # Proportionality function
  ) %>% 
  pivot_wider(names_from = PropertyType, values_from = Variable) %>% 
  drop_na() %>% 
  mutate(
    A = A / All,
    I = I / All,
    O = O / All,
    R = R / All,
    H = H / All
  ) %>% 
  select(-c("All")) %>% 
  pivot_longer(cols = c("A", "I", "O", "R", "H"), names_to = "PropertyType", values_to = "Proportion") %>% 
  mutate(PropertyType = as.factor(PropertyType)) 
  
ncf_fplot %>%   
  ggplot(mapping = aes(x = PropertyType, y = Proportion)) +
  geom_boxplot() +
  scale_y_continuous(labels = percent) +
  theme_bw()
  
```

```{r}
ncf_fplot %>% 
  group_by(PropertyType) %>% 
  summarise(Proportion = mean(Proportion)) %>% 
  ggplot(mapping = aes(x = PropertyType, y = Proportion, fill = PropertyType)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar()
```


# 3. Basis risk on the NPI

## 3.1. Binomial model

```{r Binomial model for pricing derivatives}
# Determine spot price given total number of steps n, 
# and number of upward movements u_n.
spot_price <- function(s0, u, d, n, u_n = n) 
  s0 * (1 + u) ^ (u_n) * (1 + d) ^ (n - u_n)

# Risk-free asset value at moment n.
rf_price <- function(a0, r, n) 
  a0 * (1 + r) ^ n

# Plain vanilla options payoff
dvt_call <- function(s, k) max(s - k, 0)
dvt_put <- function(s, k) max(k - s, 0)

# Binomial model to price an option on the NPI
dvt_binom <- function(h, s, u, d, r, N) {
  
  # Function that determine number of U's given the number of steps
  f_us <- function(n) {
    
    # Switch flag
    f_flag <- function() ifelse(flag == 1, flag <<- 0, flag <<- 1)
    
    # Populate sample space as a matrix 
    n_row <- 2 ^ n; n_col <- n
    us <- matrix(rep(0, n_row * n_col), nrow = n_row, ncol = n_col)
    for (j in n_col:1) {
      flag <- 1 # u
      n_perGeom <- 2 ^ (n_col - j) # Geometric trick
      k <- n_perGeom
      for (i in 1:n_row) {
        if (k == 0) {
          f_flag()
          k <- n_perGeom - 1
        } else {
          k <- k - 1
        }
        us[i, j] <- flag
      }
    }
 
    # Return number of u's per tree branch
    return(apply(us, 1, sum))
  }
  
  # Risk-neutral probability
  q <- (r - d) / (u - d)

  # Backward prop
  for (i in N:0) {
    v <- vector("double", length = 2 ^ i)  
    k <- 1
    # Right hand side of the tree (implicit value)
    if (i == N) {
      v <- sapply(spot_price(s, u, d, N, u_n = f_us(N)),h)
    # Otherwise
    } else {
      for (j in 1:(2 ^ i)) {
        v[j] <- (q * v0[k] + (1 - q) * v0[k + 1]) / (1 + r)
        k <- k + 2
      }
    }
    # Next iteration
    v0 <- v
  }
  
  # Return derivative price
  return(v0)
}

# CRR valuation
dvt_crr <- function(h, s, u, d, r, N) {
  q <- (r - d) / (u - d)
  scenarios <- choose(N, 0:N) * q ^ (0:N) * (1 - q) ^ (N:0) * sapply(
    s * (1 + u) ^ (0:N) * (1 + d) ^ (N:0), h)
  sum(scenarios) * (1 + r) ^ (-N)
}

# BSM valuation
dvt_bsm <- function(S, K, rf, Mat, sigma, Exc) {
  
  if (Mat > 0) { #BSM formula
    d1 <- (log(S / K) + (rf + 0.5 * sigma ^ 2) * Mat) / (sigma * sqrt(Mat))
    d2 <- d1 - sigma * sqrt(Mat)
    if (Exc == "Call") {
      price <- S * pnorm(d1) - K * exp(-rf * Mat) * pnorm(d2)
    } else {  # Put
      price <- K * exp(-rf * Mat) * pnorm(-d2) - S * pnorm(-d1)
    }
    
    return(price)
    
  } else { # Intrinsic value
    if (Exc == "Call") {
      dvt_call(S, K)
    } else {
      dvt_put(S, K)
    }
  }
  
} 

```

## 3.2. Price and basis risk calculations.

Risk free asset data (3 Months T-Bills)

```{r Import risk-free asset data, message=FALSE, warning=FALSE}
rf_data <- read_csv("../Source/SourceFiles/NCREIF/TB3MS.csv") %>% 
  transmute(
    Date = DATE %m+% days(-1),
    rf = TB3MS / 100
  )

rf_data %>% 
  ggplot(mapping = aes(x = Date, y = rf)) +
  geom_line() +
  scale_y_continuous(labels = percent) +
  theme_bw()
```

Joining NPI (spot) data with risk free asset to price derivatives.

**Derivatives condition**

European call options:

*Fixed inputs*

- S: NPI Index
- rf: 3 months T-Bills
- $\sigma$: EWMA with $\lambda = 0.94$.

*Parameters*

- K: to be parametrized (itm, atm, otm)
- T: 3 years (would be good to use 5 years too)

Let bring the data.

```{r Market data for option pricing}
# Function that calculates EWMA volatility given a returns vector
spot_ewma <- function(x, lambda = 0.94) {
  n <- length(x)
  v <- vector("double", n)
  v[1] <- var(x)
  for (i in 2:n) {
    v[i] <- lambda * v[i - 1] + (1 - lambda) * (x[i - 1] ^ 2)
  }
  return(sqrt(v))
}

# Market data set for pricing options
dvt_mktdata <- ncf_dataset %>% 
  filter(PropertyType == "All") %>% 
  select(Date, r) %>% 
  mutate(NPI = cumprod(1 + r) * 100) %>%
  mutate(vol = sqrt(4) * spot_ewma(r)) %>% 
  left_join(rf_data, by = c("Date")) %>% 
  mutate(rf = 4 * log(1 + rf / 4))
```

Now let's "write" 3-year derivatives each quarter from 1990-03-31 to 2017-12-31 and calculate the basis risk using at the money calls.

```{r}

# Derivatives years to maturity
ytm <- 3

# Calculate prices
lapply(
  dvt_mktdata %>% 
    filter(
      # Date >= as.Date("1985-03-31"),
      Date >= as.Date("1998-12-31"),
      Date <= as.Date("2017-03-31")
    ) %$% Date, 
  function(d) {
    dfp <- dvt_mktdata %>% 
      filter(
        Date >= d,
        Date <= d %m+% months(12 * ytm)
      ) %>% 
      mutate(
        ttm = as.numeric(max(Date) - Date) / 365, 
        strike = rep(NPI[1], 4 * ytm + 1),
        exercise = "Put",
        put = pmap_dbl(list(NPI, strike, rf, ttm, vol, exercise), dvt_bsm)
      ) %>% 
      select(Date, put) %>% 
      transmute(Date, dc = c(NA, diff(put))) 
    
    br_model <- function(dataset) {
      lm(Delta ~ dc, data = dataset)
    }
    
    port_returns %>% 
      mutate(Portfolio = map(Portfolio, function(x) {
        x %>% 
          left_join(dfp, by = c("Date")) %>% 
          select(Date, dc, Delta) %>% 
          drop_na()
      })) %>% 
      mutate(
        linear_model = map(Portfolio, br_model),
        summary = map(linear_model, function(l) {
          x <- l$model[, 2, drop = TRUE]
          y <- l$model[, 1, drop = TRUE]
          e <- l$residuals
          n <- length(x)
          sxx <- sqrt(sum(x ^ 2) - (sum(x) ^ 2) / n)
          s <- sqrt(sum(e ^ 2) / (n - 2))
          return(
            tibble(
              beta = l$coefficients[2],
              sigma = s,
              t.statistic = beta / (s / sxx),
              p.value = pt(-abs(t.statistic), n - 2, 0.025) + (1 - pt(abs(t.statistic), n - 2, 0.025))
            )
          )
        })
      ) %>% 
      select(key, summary) %>% 
      unnest(summary) %>% 
      write_csv(str_c("basis_risk_put/", dfp[4 * ytm + 1, 1, drop = TRUE], ".csv"))
    
    return(d)
  })

```

```{r}
# To verify
port_returns %>% 
      mutate(Portfolio = map(Portfolio, function(x) {
        x %>% 
          left_join(dfp, by = c("Date")) %>% 
          drop_na()
      })) %>% 
      mutate(
        linear_model = map(Portfolio, br_model),
        glance_model = map(linear_model, glance)
      ) %>% 
      select(key, glance_model) %>% 
      unnest(glance_model) %>% View()
```


Not always basis risk is hedged. Coefficient is not significant always. Can we improve the hedging using basket options?



